The prime of their life - the halflife of acting talent

### 2017-09-04 01_halflife-of-acting-talent/article.Rmd
stuff

.finding {
  background-color:#000;
  height: 125px;
  color: #FFF;
  display: flex;
  border-radius: 25px;
  align-items: center;
  text-align:center;
}

.finding .sidebar { 
  padding-top:15px;
  width: 30%;
  font-weight: bold;
  font-size: 18px;
}

.finding .body {
  width: 70%;
  text-align:left;
  font-size: 16px;
  padding-right:10px
}

@@

#the-half-list-of-acting-talent img {
  height: 200px; 
  display:block; 
  margin: 0 auto;
}

@@

# Finding function

create_finding = function(finding_text) {

  html = 
glue::glue("<br><br><div class='finding'>

<div class='sidebar'>
  <img src='https://github.com/TMBish/lab_posts/blob/master/00_utilities/lightbulb.png?raw=true'>
  <span> Finding!</span>
</div>

<div class='body'> 
  <span> {finding_text} </span> 
</div>

</div><br><br>")

  return(html)

@@

So generally this is looking pretty, pretty good. Things get dicey with unisex names like Sandy; however, the package must know that most Sandys born in 1937 were female so we'll have to trust it here.

@@

    base_chart = 
      highchart() %>%
      hc_chart(type = "column") %>%
      hc_add_theme(tmbish) %>%
      hc_xAxis(categories = data$role_type) %>%
      hc_plotOptions(
        column = list(
          stacking = 'percent',
          dataLabels = list(
            enabled = TRUE,
            color = "#FFF",
            style = list(textOutline = NULL, fontSize = "14px"),
            formatter = JS("function(){ return(Math.round(this.percentage) + '%')}")
          )
        )
      ) %>%
      hc_title(text = ifelse(metric=="films", "By Volume of Films", "By Distinct Individuals")) %>%
      hc_legend(align = "center", verticalAlign = "bottom", layout = "horizontal")

@@

    if (metric != "films") { base_chart  = base_chart %>% hc_legend(enabled=FALSE)}

@@

gender_prop_finding = create_finding("Approximately 80% of all directors are male, and 85% of popular films are directed by a man.")

gender_props %>%
  hw_grid(ncol = 1, rowheight = "400px") %>%
  browsable()

@@

So we'll treat directors as a single group that, statistically speaking, are very much likely to be males.

@@

# Create chart
age_dist_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(
    animation = list(duration = 2000)
  ) %>%
  hc_xAxis(
    categories = chart_data$age_at_production,
    title = list(text = "Age at Production")
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    labels = list(formatter = JS("function(){return(Math.round(this.value * 1000) / 10 + '%')}"))
  ) %>%
  hc_title(text = "Age at Production") %>%
  hc_subtitle(text = "Comparing the distribution of film professionals for age at the time of film productions") %>%
  hc_add_series(
    name = "Director", 
    type = "spline",
    visible= FALSE,
    data = chart_data$director,
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Male Actor",
    type = "spline",
    visible= FALSE,
    data = chart_data$male_actor,
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Female Actor",
    type = "spline",
    data = chart_data$female_actor,
    zIndex = -10,
    marker = list(enabled = FALSE)
  ) %>%
  hc_tooltip(
    formatter = JS("function(){return('<b> Series: </b>' + point.key + '<br> <b> Age: </b>' + this.x + '<b> Value: </b>' + this.y )}")
  )

age_atprod_finding = create_finding("64% of female roles are given to actors under the age of 40 compared with 42% of male roles. Only 28% of popular films are directed by a someone under 40.")

@@

@@

  highchart() %>%
    hc_add_theme(tmbish) %>%
    hc_chart(
      animation = list(duration = 2000), shadow = FALSE
    ) %>%
    hc_xAxis(
      categories = actordata$age_bucket,
      title = list(text = "Age at Production")
    ) %>%
    hc_yAxis(labels = list(formatter = JS("function(){return(this.value*100 + '%')}"))) %>%
    hc_title(text = actor) %>%
    hc_add_series(
      name = "Performance Index",
      type = "spline",
      data = actordata$index
    ) %>%
    hc_legend(enabled = FALSE)

}) %>% hw_grid(ncol = 3, rowheight = 350) %>% browsable()

@@

volume_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(height = 300, spacingRight = 125) %>%
  hc_xAxis(categories = rating_volumes$age_at_production, opposite = TRUE) %>%
  hc_yAxis(
    title = list(text = "Total Films in Dataset"),
    labels = list(formatter = JS("function(){return(-1 * this.value)}"))    
    ) %>%
  hc_title(text = "") %>%
  hc_add_series(
    name = "Volumes",
    type = "area",
    data = rating_volumes$films,
    marker = list(enabled = FALSE)
  ) %>% 
  hc_legend(enabled = FALSE)

rating_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(
    animation = list(duration = 2000)
  ) %>%
  hc_xAxis(
    categories = rating_scores$age_at_production,
    title = list(text = "Age at Production"),
    plotBands = list(
       list(
         from = 27-15, 
         to = 35-15,
         color = "rgba(255,106,92,0.2)",
         label = list(text = "Acting prime: 27-35!", style=list(fontFamily="Arvo",fontWeight="bold", fontSize=10), verticalAlign="middle"))
    )
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    min = -0.4,
    labels = list(
      formatter = JS("function(){return(this.value*100 + '%')}")
    )
  ) %>%
  hc_title(text = "Female actors") %>%
  hc_subtitle(text = "Comparing male actors, (male) directors and female actors") %>%
  hc_add_series(
    name = "Director",
    type = "scatter",
    data = rating_scores$director,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Male Actor",
    type = "scatter",
    data = rating_scores$male_actor,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Female Actor",
    type = "scatter",
    data = rating_scores$female_actor,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Trend",
    type = "spline",
    data = predictions,
    color = "#000000",
    marker = list(enabled = FALSE)
  )

@@

While it appears that actors and directors have their prime between the ages of 27-35 I don't think it's fair to say this is **caused** by a natural skill trajectory. Some caveats that could potentially explain this trend without our skills and abilities narrative:

* Most films typically have non-elderly main characters; limiting the ability for an older, but still skillful, actor to be cast in these roles
* Seperating an Actor's role **choice** vs their **performance** given they chose a particular role highlights the problem we see with De Niro above. It's quite possible that after career success and achievements actor's and directors prioritise different things (eg. *$Cash Money$*) in the twiglights of their careers.

>>>>

.finding {
  background-color:#414141;
  height: 125px;
  color: #FFF;
  display: flex;
  align-items: center;
  text-align:center;
}

.finding .sidebar { 
  padding-top:15px;
  width: 20%;
  font-weight: bold;
  font-size: 18px;
}

.finding .body {
  width: 75%;
  text-align:left;
  font-size: 16px;
  padding-right:10px
}

@@

#snatch img {
  height: 200px; 
  display:block; 
  margin: 0 auto;
}

#no-country-for-old-men img {
  height: 200px; 
  display:block; 
  margin: 0 auto;
}

#the-departed img {
  height: 200px; 
  display:block; 
  margin: 0 auto;
}

@@

# Finding function
source("https://raw.githubusercontent.com/TMBish/lab_posts/master/00_utilities/create-finding-html.R")
```

@@

So generally this is looking pretty, pretty good. Things get dicey with unisex names like Sandy; however, the package must know that Sandy was more commonly used for females in the 1930s, which is consistent with my [1 data other point.](https://www.youtube.com/watch?v=cxbMk-8701M)

@@

    base_chart = 
      highchart() %>%
      hc_chart(type = "column", shadow=FALSE) %>%
      hc_add_theme(tmbish) %>%
      hc_xAxis(categories = data$role_type) %>%
      hc_yAxis(title = list(text = ifelse(metric=="films", "Percentage of Total Roles", "Percentage of Distinct Individuals"))) %>%
      hc_plotOptions(
        column = list(
          stacking = 'percent',
          dataLabels = list(
            enabled = TRUE,
            color = "#FFF",
            backgroundColor = NULL,
            style = list(textOutline = NULL, fontSize = "14px"),
            formatter = JS("function(){ return(Math.round(this.percentage) + '%')}")
          )
        )
      ) %>%
      hc_title(text = ifelse(metric=="films", "By Volume of Roles", "By Distinct Individuals")) %>%
      hc_legend(enabled=FALSE) %>%
      hc_tooltip(
        formatter = JS("function(){return('<b>Gender: </b>' + this.series.name + '<br> <b>Proportion: </b>' + Math.round(this.percentage) + '%')}")
      )
      # hc_legend(align = "center", verticalAlign = "bottom", layout = "horizontal")

@@

@@

gender_prop_finding = create_finding("Approximately 85% of all directors are male, and almost 90% of popular films are directed by a man.")

gender_props %>%
  hw_grid(ncol = 2, rowheight = "400px") %>%
  browsable()

@@

So we'll treat directors as a single group that, statistically speaking, are very likely to be male.

@@

# Create chart
age_dist_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(
    animation = list(duration = 2000)
  ) %>%
  hc_xAxis(
    categories = chart_data$age_at_production,
    title = list(text = "Age at Production")
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    labels = list(formatter = JS("function(){return(Math.round(this.value * 1000) / 10 + '%')}"))
  ) %>%
  hc_title(text = "Age at Production") %>%
  hc_subtitle(text = "Contrasting the distribution of ages at production across gender and role") %>%
  hc_add_series(
    name = "Director", 
    type = "spline",
    visible= FALSE,
    data = chart_data$director,
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Male Actor",
    type = "spline",
    visible= FALSE,
    data = chart_data$male_actor,
    marker = list(enabled = FALSE)
  ) %>%
  hc_add_series(
    name = "Female Actor",
    type = "spline",
    data = chart_data$female_actor,
    zIndex = -10,
    marker = list(enabled = FALSE)
  ) %>%
  hc_tooltip(
        formatter = JS("function(){return('<b> Series: </b>' + this.series.name + '<br> <b>Age: </b>' + this.x + '<br> <b>Value: </b>' + Math.round(this.y*1000) / 10 +'%' )}")
  )

@@

```{r results='asis'}
age_atprod_finding = create_finding("64% of female roles are given to actors under the age of 40 compared with 42% of male roles. Only 28% of popular films are directed by a someone under 40.")

age_atprod_finding

@@

  highchart() %>%
    hc_add_theme(tmbish) %>%
    hc_chart(
      animation = list(duration = 2000), shadow = FALSE
    ) %>%
    hc_xAxis(
      categories = actordata$age_bucket,
      title = list(text = "Age at Production")
    ) %>%
    hc_yAxis(labels = list(formatter = JS("function(){return(this.value*100 + '%')}"))) %>%
    hc_title(text = actor) %>%
    hc_add_series(
      name = "Performance Index",
      type = "spline",
      data = actordata$index
    ) %>%
    hc_legend(enabled = FALSE) %>%
    hc_tooltip(
         formatter = JS("function(){return('<b>Tomato Index: </b>' + Math.round(this.y*1000) / 10 +'%' )}")
    )

}) %>% hw_grid(ncol = 3, rowheight = 250) %>% browsable()

@@

volume_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(height = 300, spacingRight = 125, shadow=FALSE) %>%
  hc_xAxis(categories = rating_volumes$age_at_production, opposite = TRUE) %>%
  hc_yAxis(
    title = list(text = "Total Films in Dataset"),
    labels = list(formatter = JS("function(){return(-1 * this.value)}"))    
    ) %>%
  hc_title(text = "") %>%
  hc_add_series(
    name = "Volumes",
    type = "area",
    fillOpacity = 0.25,
    data = rating_volumes$films,
    marker = list(enabled = FALSE)
  ) %>% 
  hc_legend(enabled = FALSE) %>%
    hc_tooltip(
       formatter = JS("function(){return('<b>Total Performances At This Age: </b>' + -1*this.y)}")
  )

rating_chart = 
  highchart() %>%
  hc_add_theme(tmbish) %>%
  hc_chart(
    animation = list(duration = 2000),
    shadow = FALSE
  ) %>%
  hc_xAxis(
    categories = rating_scores$age_at_production,
    title = list(text = "Age at Production"),
    plotBands = list(
       list(
         from = 27-15, 
         to = 35-15,
         color = "rgba(255,106,92,0.2)",
         label = list(text = "Acting prime: 27-35!", style=list(fontFamily="Arvo",fontWeight="bold", fontSize=10), verticalAlign="middle"))
    )
  ) %>%
  hc_yAxis(
    title = list(text = "Probablity Density"),
    min = -0.4,
    labels = list(
      formatter = JS("function(){return(this.value*100 + '%')}")
    )
  ) %>%
  hc_title(text = "The Prime of Your Life") %>%
  hc_subtitle(text = "Comparing male actors, (male) directors and female actors") %>%
  hc_add_series(
    name = "Director",
    type = "scatter",
    data = rating_scores$director,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Male Actor",
    type = "scatter",
    data = rating_scores$male_actor,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Female Actor",
    type = "scatter",
    data = rating_scores$female_actor,
    marker = list(radius = 3, symbol = "circle")
  ) %>%
  hc_add_series(
    name = "Trend",
    type = "spline",
    data = predictions,
    color = "#000000",
    marker = list(enabled = FALSE)
  ) %>%
  hc_tooltip(
        formatter = JS("function(){return('<b>Role Type: </b>' + this.series.name + '<br> <b>Age: </b>' + this.x + '<br> <b>Average Tomato Index: </b>' + Math.round(this.y*1000) / 10 +'%' )}")
  )

@@

While it appears that actors and directors have their prime between the ages of 27-35 we can't say with confidence that this is **caused** by a natural skill trajectory. Some caveats that could potentially explain this trend without our skills and abilities narrative:

* Most films (and by extension good films) typically have non-elderly main characters; limiting the ability for an older, but still skillful, actor to be cast in these roles
* Seperating an actor's role **choice** vs their **performance** given they chose a particular role highlights the problem we see with De Niro above. It's quite possible that after career success and personal skill based accolades, actor and directors prioritise different things (eg. **$ Cash Money $**) in the twiglight of their careers.
